---
title: Rollbar
author: Freddy Drennan
date: '2018-02-11'
slug: rollbar
categories:
  - errors
  - rollbar
tags: []
---

I've been working on creating my own version of RSConnect by hand. I think it's a bit awful that anyone would have to shell out 25K a year for software. 

With that said, part of the process was trying to figure out how to deal with errors that occur within my Docker images. The solution I found was [Rollbar](http://rollbar.com). It's super easy to get started in R.

Log in and create an account. The first image you'll see is this:


```{r ,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/rollbar1.png")
 grid.raster(img)
```


Pick a new project and name it. 
```{r ,echo=FALSE}
img <- readPNG("images/rollbar2.png")
 grid.raster(img)
```

Then, yea.. I'm sorry... select __python__.
```{r ,echo=FALSE}
img <- readPNG("images/rollbar3.png")
 grid.raster(img)
```

Then, copy the secure key. Mine is now deleted, so you can't use mine. 
```{r ,echo=FALSE}
img <- readPNG("images/rollbar4.png")
 grid.raster(img)
```


I already have rollbar installed on my computer. You need to run ```install.packages("rollbar")``` first. Also, given that this is not a package that runs frequenly, I prefer to call functions explicity instead of loading the packge. 

```{r}
rollbar::rollbar.configure(access_token="d576994698c248faa7c990e16fee456a", env="production")
rollbar::rollbar.error('Days > 30 in weather_plot API')
```

You'll see that the code ran well.
```{r, echo=FALSE}
img <- readPNG("images/rollbar5.png")
 grid.raster(img)
```

You can go under settings and get notifications to Slack as well!
```{r ,echo=FALSE}
img <- readPNG("images/rollbar6.png")
 grid.raster(img)
```


Next try this!

Create this file and name it rollbar.R
```
library(rollbar)
rollbar.configure(access_token="d1ed4487e4d7403b82c992448f3c5602", 
                           env="production",
                           root="/Users/freddydrennan/Documents/programming/rProjects/fdrennan")

rollbar.attach()
print(a)

```

Then run ```Rscript rollbar.R```

and you'll get errors sent for general exceptions!