---
output:
  html_document: default
  html_notebook: default
---


 <meta charset="utf-8" />
<link rel="stylesheet" href="~/~/css/font-awesome.min.css">


```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
cat("Last push: ", as.character(Sys.time()))
```




<!-- Footer -->
<header>
<!-- <div class="container"> -->
<div class="row">
<div class="col-lg-8 col-lg-offset-2 col-md-10 col-md-offset-1">
<ul class="list-inline text-center">
<li>
<a href="mailto:drennanfreddy@gmail.com?Subject=Hello">
<span class="fa-stack fa-lg">
<i class="fa fa-circle fa-stack-2x"></i>
<i class="fa fa-envelope fa-stack-1x fa-inverse"></i>
</span>
</a>
</li>
<li>
<a href="https://www.linkedin.com/in/freddy-drennan-4a8b79aa">
<span class="fa-stack fa-lg">
<i class="fa fa-circle fa-stack-2x"></i>
<i class="fa fa-linkedin fa-stack-1x fa-inverse"></i>
</span>
</a>
</li>
<li>
<a href="http://facebook.com/freddy.drennan">
<span class="fa-stack fa-lg">
<i class="fa fa-circle fa-stack-2x"></i>
<i class="fa fa-facebook fa-stack-1x fa-inverse"></i>
</span>
</a>
</li> 
<li>
<a href="https://github.com/fdrennan">
<span class="fa-stack fa-lg">
<i class="fa fa-circle fa-stack-2x"></i>
<i class="fa fa-github fa-stack-1x fa-inverse"></i>
</span>
</a>
</li>
</ul>
</div>
</div>
</div>
</header>


<center>
# Economic Indicators
</center>


```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
threeYears = Sys.Date() - 365*10
threeYears = paste0(threeYears, "::")
```

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
library(quantmod)
library(Quandl)
# library(data.table)
invisible(getSymbols(c("GDP", "MKTGDP1WA646NWDB"), src = "FRED"))
```

 US GDP
```{r}
# US GDP
invisible(chartSeries(GDP[threeYears], name = "GDP"))
# WORLD GDP
```


 World GDP
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(chartSeries(MKTGDP1WA646NWDB["2005::"], name = "World GDP"))
```

 Real GDP Percent Change
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
# Real GDP, percent change from preceding period
invisible(getSymbols("A191RL1Q225SBEA", src = "FRED"))

chartSeries(A191RL1Q225SBEA[threeYears], name = "Read GDP: Percent Change")
```

 GDP Price Deflator
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("GDPDEF", "CPIAUCSL"), src = "FRED"))
chartSeries(GDPDEF[threeYears], name = "GDP Price Deflator")
```

 CPI
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
chartSeries(CPIAUCSL[threeYears], name = "Consumer Price Index for All Consumers: All Items")
```



<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- library(fdRen) -->
<!-- bonds=fdRen::bondTable() -->
<!-- bills=as.xts(bonds$Bills[,5], order.by = bonds$Bills[,1]) -->
<!-- billsChange=as.xts(bonds$Bills[,4], order.by = bonds$Bills[,1]) -->
<!-- notes=as.xts(bonds$NotesBonds[,6], order.by = bonds$NotesBonds[,1]) -->
<!-- notesChange=as.xts(bonds$NotesBonds[,5], order.by = bonds$NotesBonds[,1]) -->
<!-- tips=as.xts(bonds$tips[,7], order.by = bonds$tips[,1]) -->
<!-- ``` -->

<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- chartSeries(bills, name = "Treasury Bills Yeild Curve") -->
<!-- ``` -->

<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- chartSeries(billsChange, name = "Treasury Bills Change") -->
<!-- ``` -->


<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- chartSeries(notes, name = "Treasury Notes Yeild Curve") -->
<!-- ``` -->

<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- chartSeries(notesChange, name = "Treasury Notes Change") -->
<!-- ``` -->

<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- chartSeries(tips, name = "Tips Yeild Curve") -->
<!-- ``` -->

 Shiller PE
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
library(fdRen)
dat = urlTables("http://www.multpl.com/shiller-pe/table")
dat=dat[[1]]

colnames(dat) = c("Date", "Value")
dat$Date = as.Date(dat$Date,
                   format ="%b %d, %Y")

dat=as.xts(dat[,2], order.by = dat[,1])
chartSeries(dat, name = "Shiller PE")
```

 Shiller Home Price Index
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols("CSUSHPINSA", src="FRED"))
chartSeries(CSUSHPINSA[threeYears], name = "S&P/Case-Shiller U.S. National Home Price Index")
```


```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10, eval = FALSE}


library(fdRen)
tickers = urlTables("https://en.wikipedia.org/wiki/List_of_S%26P_500_companies", fill = TRUE)
tickers = tickers[[1]]
tickers = tickers
# ticks = c(tickers[,1], "SPY", "GLD", "SHV", "BND", "TSLA", "SNAP")


stuff = c("Market Capitalization", "P/E Ratio")
theData = matrix(0, ncol = (length(stuff)+1), nrow = 1)
for(i in 1:length(tickers[,1])) {
     theData = rbind(as.matrix(getQuote(tickers[i,1], what = yahooQF(c("Market Capitalization", "P/E Ratio")))), theData)
}

library(dplyr)
# pem=as.data.frame(readRDS("pemarket.Rds"))
pem = as.data.frame(theData)
pem$`Market Capitalization` = sub("B", "", pem$`Market Capitalization`)
pem$`Market Capitalization` = as.numeric(as.character(pem$`Market Capitalization`))
pem$`P/E Ratio` = as.numeric(as.character(pem$`P/E Ratio`))
pem=pem[complete.cases(pem),]
hist(pem$`Market Capitalization`, breaks = dim(pem)[1], ylab = "Frequency", xlab = "Market Cap in Billions", main = "S&P Market Cap")
pem$ticker = rownames(pem)
pem = pem[,c(4,3,2)]
colnames(pem) = c("ticker", "pe", "mc")
pem=arrange(pem, pe)
pem = pem[-1,]

pem$propMc = (pem$mc/sum(pem$mc))
pem$wape = pem$propMc*pem$pe
pem$avgwape = pem$wape/sum(pem$wape)

plot(pem$mc, pem$wape/sum(pem$wape), ylab = "Weighted Average PE Contribution", xlab = "Market Cap")
plot(pem$mc, pem$wape/sum(pem$wape), ylab = "Weighted Average PE Contribution", 
                                     xlab = "Market Cap", ylim = c(0, .04))
hist(pem$pe, breaks = 500, xlim = c(0, quantile(pem$pe, .95)), xlab = "P/E Ratio", ylab = "Frequency", main = "S&P 500 P/E Ratios")
hist(pem$pe, breaks = 300, xlab = "P/E Ratio", ylab = "Frequency", main = "S&P 500 P/E Ratios")

pem=arrange(pem, rev(avgwape))

pemNone = pem[-c(1:5),c(1:3)]
http://fdrennan.net/pages/pages2/weather.html
pemNone$propMc = (pemNone$mc/sum(pemNone$mc))
pemNone$wape = pemNone$propMc*pemNone$pe
pemNone$avgwape = pemNone$wape/sum(pemNone$wape)

cat("The Weighted Average PE is: ", sum(pemNone$wape))
cat("The Average PE is: ", mean(pemNone$pe))

colnames(pem) = c("Ticker", "PE Ratio", "Market Cap", "MC Proportion", "Weighted PE", "PE Cont")
pem[,2:6]= round(pem[,2:6], 5)
DT::datatable(pem)
write.csv(pem, "/home/fdrennan/Documents/r/website/pem.csv")
library(RCurl)
invisible(ftpUpload("/home/fdrennan/Documents/r/website/pem.csv", 
"ftp://fdrennan.net/public_html/pages/pages2/datasets/pem.csv", 
    userpwd="fdrennan:Thirdday1!"))
```
Download the above table [here](http://fdrennan.net/pages/pages2/datasets/pem.csv)


<!-- ```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10} -->
<!-- invisible(getSymbols(c("LFWA64TTUSM647S"), src = "FRED")) -->
<!-- chartSeries(LFWA64TTUSM647S, name = "Working Age Population: Ages 15-64") -->
<!-- ``` -->

 Civilian Unemployment Rate
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols("UNRATE", src = "FRED"))
chartSeries(UNRATE[threeYears], name = "Civilian Unemployment Rate",
            show.grid = TRUE)
```



The PMI (R) is a composite index based on the diffusion indexes of five of the indexes
with equal weights: New Orders (seasonally adjusted), Production (seasonally adjusted),
Employment (seasonally adjusted), Supplier Deliveries (seasonally adjusted), and Inventories. A value above 50 suggests positive outlook and below 50 the opposite. 

 Manufacturing Confidence Index
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
library(Quandl)
PMI = Quandl("ISM/MAN_PMI")
PMI = as.xts(PMI[,2],order.by = PMI[,1])
chartSeries(PMI[threeYears], name = "PMI")
```

 Housing Starts
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("HOUST", "CSUSHPINSA"), src = "FRED"))
chartSeries(HOUST[threeYears], name = "Housing Starts")
```


 Hiring 
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("JTSHIL"), src = "FRED"))
chartSeries(JTSHIL["2005-01-01::"], name = "Hiring")
```

 Capacity Utilization
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("TCU"), src = "FRED"))
chartSeries(TCU, name = "Capacity Utilization")
```

 Stock Market Capitalization to GDP for US
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("DDDM01USA156NWDB"), src = "FRED"))
chartSeries(DDDM01USA156NWDB, name = "Stock Market Capitalization to GDP for United States")
```
Interesting document [here](https://fredblog.stlouisfed.org/2016/11/the-continuous-rise-of-the-stock-market/?utm_source=series_page&utm_medium=related_content&utm_term=related_resources&utm_campaign=fredblog) about the continuous rise of the stock market. 

 Stock Market Total Value Traded to GDP
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("DDDM02USA156NWDB"), src = "FRED"))
chartSeries(DDDM02USA156NWDB, name = "Stock Market Total Value Traded to GDP for United States")
```

 
 10-Year Treasury Constant Maturity Rate
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("DGS10"), src = "FRED"))
chartSeries(DGS10["1990-01-01::"], name = "10-Year Treasury Constant Maturity Rate")
```



 Fed Funds and Reverse RPs with Banks, All Commercial Banks 
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("FRPACBW027SBOG"), src = "FRED"))
chartSeries(FRPACBW027SBOG["1988-01-01::"], name = "Fed Funds and Reverse RPs with Banks, All Commercial Banks")
```


 Effective Federal Funds Rate
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("FEDFUNDS"), src = "FRED"))
chartSeries(FEDFUNDS, name = "Effective Federal Funds Rate")
```

 Interest Rates, Discount Rate for United States
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.height=6, fig.width=10}
invisible(getSymbols(c("INTDSRUSM193N"), src = "FREhttp://fdrennan.net/pages/pages2/weather.htmlD"))
chartSeries(INTDSRUSM193N, name = "Interest Rates, Discount Rate for United States")
```
