---
title: SQL for LinkedIn
author: Freddy Drennan
date: January 4, 2019
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Here's the deal. You have access to a database running on my website drenr.com.

You can read/write/destroy anything in the database linkedin. Preferably you don't delete everything. Mistakes are ok. It's Postgres. 

Enjoy playing and learning. Don't be a butt... intentionally.

[Official Docs](https://www.postgresql.org/docs/10/sql-commands.html)

[R Docs](https://db.rstudio.com/databases/postgresql/)

```{r}
library(RPostgreSQL)
library(tidyverse)
```

### Connecting
```{r}

con <- dbConnect(PostgreSQL(),
                 dbname   = 'linkedin',
                 host     = 'drenr.com',
                 port     = 5432,
                 user     = "linkedin_user",
                 password = "password")

```

### List of tables that exist currently
```{r}
dbListTables(con)
```

### Grab an entire table
```{r}
dbReadTable(con, "p_weather") %>% head
```

### Send a query, do dplyr server side, then download. 
```{r}
query <- '
SELECT *
FROM mtcars
WHERE 1 = 1
'

results <- tbl(con, sql(query))

results
```


```{r}
results <- 
  results %>% 
   # server side aggregation
  filter(hp == 110) %>% 
  group_by(am) %>% 
  count 

# Downloading the data
results = as.data.frame(results)
results
```


### Slightly more complicated query.
```{r}
query <- '
SELECT 
  COUNT(*),
  year,
  month,
  day
FROM 
  p_flights
GROUP BY 
  year, 
  month,
  day
ORDER BY year, month, day
'

results <- tbl(con, sql(query)) %>% as.data.frame()


results %>% 
  mutate(date = as.Date(paste0(year, "-", month, "-", day))) %>% 
  ggplot() +
  aes(x = date, y = count) +
  geom_line()
```

### Inserting a table

Choose a name that isnt taken above for name field. Insert your dataframe in the value field. If the table exists, the data will be appended assuming it is the same structure. 
```{r}
# dbWriteTable(conn = con,
#              name = "somedataframe",
#              value = mtcars,
#              append = TRUE,
#              row.names = FALSE)
```

### DISCONNECT WHEN YOU ARE DONE
```{r}
dbDisconnect(con)
```
