---
title: "False Position"
author: "Freddy Drennan"
date: "11/21/2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: cerulean
    highlight: haddock
    fig_align: center
---

A way to avoid the pathology of the secant method is to ensure that the root is bracketed between the two starting values and remains between the successive pairs. When this is done, the method is known as linear interpolation, or more often, as the method of false position. This technique is similar to bisection except the next iterate is taken at the intersection of a line between the pair of x-values and the x-axis, rather than at the midpoint. Doing so gives a faster convergence than does bisection, but at the expense of a more complicated algorithm. 


```{r}
library(reticulate)
library(zeallot)
use_python('/Users/digitalfirstmedia/Documents/R/numerical_analysis/numerical_analysis/bin/python3')
```

To determine a root of $f(x) = 0$ given values of $x_0$ and $x_1$ that bracket a root, that is, $f(x_0)$ and $f(x_1)$ are of opposite sign,

```
REPEAT
  Set x_2 = x_0 - f(x_0) * (x_0 - x_1) / [f(x_0) - f(x_1)]
  IF f(x_2) of opposite sign to f(x_0)
    Set x_1 = x_2
  ELSE Set x_0 = x_2
  ENDIF
UNTIL |f(x_2)| < tolerance value
```

## R
```{r}

false_position <- function(f, x_0, x_1, verbose = TRUE, tolerance = 1e-10) {

  repeat {
    x_2 = x_0 - f(x_0) * (x_0 - x_1) / (f(x_0) - f(x_1))
    if(verbose) print(x_2)
    if(sign(f(x_2) != sign(f(x_0))))
      x_1 = x_2
    else x_0 = x_2
    if(abs(f(x_2)) < tolerance)
      break
  }
    
  return(x_2)
  
}

f = function(x) {
  x^2 - 16
}

x_0 = 5
x_1 = 1
tolerance = 1e-10

false_position(f, x_0, x_1)
```


## Python
```{python}
import numpy as np

def false_position(f, x_0, x_1, verbose = True, tolerance = 1e-10):
  while True:
    x_2 = x_0 - f(x_0) * (x_0 - x_1) / (f(x_0) - f(x_1))
    if verbose: 
      print(x_2)
    if np.sign(f(x_2)) != np.sign(f(x_0)):
      x_1 = x_2
    else:
      x_0 = x_2
    if np.abs(f(x_2)) < tolerance:
      break
  return x_2

def f(x):
  return x ** 2 - 16
  
x_0 = 5
x_1 = 1
tolerance = 1e-10

print(false_position(f, x_0, x_1))
```