---
title: "The Secant Method"
author: "Freddy Drennan"
date: "11/20/2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: cerulean
    highlight: haddock
    fig_align: center
---

Although the bisection (interval halving) method is easy to compute and has simple error analysis, it is not very efficient. For most functions, we can improve the speed at which the root is approached through a different scheme. Almost every function can be approximated by a straight line over a small interval. We begin from a value - say, $x_0$ - that is the to a root, $r$.

```{r}
library(reticulate)
library(zeallot)
use_python('/Users/digitalfirstmedia/Documents/R/numerical_analysis/numerical_analysis/bin/python3')

```

To determine the root of $f(x)=0$, given two values, $x_0$ and $x_1$, that are near the root.
```
IF |f(x_0)| < |f(x_1)|:
  Swap x_0 with x1
REPEAT
  Set x_2 = x_0 - f(x_0)*(x_0 - x_1)/[f(x_0) - f(x_1)]
  Set x_0 = x_1
  Set x_1 = x_2
Until |f(x_2)| < tolerance value
```
## R
```{r}

secant_method <- function(f, x_0, x_1, verbose = TRUE, tolerance = 1e-10) {
  if (abs(f(x_0) < abs(f(x_1)))) 
    # swap variables
    c(x_1, x_0) %<-% c(x_0, x_1)
  
  repeat {
    x_2 = x_0 - f(x_0)*(x_0 - x_1)/(f(x_0) - f(x_1))
    if (verbose) print(x_2)
    c(x_0, x_1) %<-% c(x_1, x_2)
    if(abs(f(x_2)) < tolerance) break
  }
    
  return(x_2)
  
}

f = function(x) {
  x^2 - 16
}

x_0 = 5000
x_1 = 10
tolerance = 1e-10

secant_method(f, x_0, x_1)
```

## Python
```{python}
import numpy as np

def secant_method(f, x_0, x_1, verbose = True, tolerance = 1e-10):
  if np.abs(f(x_0)) < np.abs(f(x_1)):
    x_1, x_0 = x_0, x_1
    
  while True:
    x_2 = x_0 - f(x_0)*(x_0 - x_1)/(f(x_0) - f(x_1))
    if verbose:
      print(x_2)
      
    x_0, x_1 = x_1, x_2
    if np.abs(f(x_2)) < tolerance:
      break
  
  return x_2
    
def f(x):
  return x ** 2 - 16
  
x_0 = 5000
x_1 = 10
tolerance = 1e-10

print(secant_method(f, x_0, x_1))
  
  
```