---
title: "Mullers Method"
author: "Freddy Drennan"
date: "11/21/2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: cerulean
    highlight: haddock
    fig_align: center
---

Another root finding method. 

```{r}
rm(list  = ls())
library(reticulate)
library(zeallot)
use_python('/Library/Frameworks/Python.framework/Versions/3.5/bin/python3')
```

Given the points $x_2$, $x_0$, $x_1$ in increasing value,

+ Evaluate the corresponding function values $f_2$, $f_0$, $f_1$.
+ Find the coefficients of the parabola determined by the three points.
+ Compute the two roots of the parabolic equation.
+ Choose the root closest to $x_0$ and label it $x_r$.
+ IF $x_r \gt x_0$ THEN rearrange $x_0$, $x_r$, $x_1$ into $x_2$, $x_0$, $x_1$.
  + ELSE rearrange $x_2$, $x_r$, $x_0$ into $x_2$, $x_0$, $x_1$.
+ IF $|f(x_r)| < \text{FTOL}$, $\text{THEN}$ return $(x_r)$.
  + ELSE go to 1
  
```{r}
f = function(x) {
  x^3 + 4*x^2 - 3
}

curve(f(x), from = -3, to = 3)
abline(h = 0)
```

## R
```{r}

mullers_method = function(f, lower, middle, upper, verbose = TRUE) {
  repeat {
    
    h_1 = upper - middle
    h_2 = middle - lower
    
    gamma = h_2/h_1
    
    a = (gamma*f(upper) - f(middle)*(1 + gamma) + f(lower))/(gamma*h_1^2*(1+gamma))
    b = (f(upper) - f(middle) - a*h_1^2)/h_1
    c = f(middle)
    
    roots = middle - (2*c)/(b + c(-1, 1)*sqrt(b^2 - 4*a*c))
    x_r = roots[abs(middle - roots) == min(abs(middle - roots))]
    
    if (verbose) print(x_r)
    
    if (x_r > middle) {
      c(lower, middle, upper) %<-% c(middle, x_r, upper)
    } else {
      c(lower, middle, upper) %<-% c(lower, x_r, middle)
    }
    
    if(abs(f(x_r)) < 1e-10) 
      return(x_r)
  }
}


result_1 = mullers_method(f, 0, .5, 1)
result_2 = mullers_method(f, -5, -2, -1)

curve(f(x), from = -2, to = 2)
abline(h = 0)
points(c(result_1, result_2), c(f(result_1), f(result_2)))
```

## Python
```{python}
import numpy as np

def mullers_method(f, lower, middle, upper, verbose = True):
  
  while True:
    
    h_1 = upper - middle
    h_2 = middle - lower
    
    gamma = h_2/h_1
    
    a = (gamma*f(upper) - f(middle)*(1 + gamma) + f(lower))/(gamma*(h_1**2)*(1+gamma))
    b = (f(upper) - f(middle) - a*h_1**2)/h_1
    c = f(middle)

    roots = middle - (2*c)/(b + np.array([-1, 1])*np.sqrt(b**2 - 4*a*c))
    
    x_r = roots[np.abs(middle - roots) == np.min(np.abs(middle - roots))]

    if verbose:
      print(x_r)

    if x_r > middle:
      lower, middle, upper = middle, x_r, upper
    else:
      lower, middle, upper = lower, x_r, middle


    if np.abs(f(x_r)) < 1e-10:
      return x_r
      
def func(x):
  return x**3 + 4*x**2 - 3


result_1 = mullers_method(func, 0, .5, 1)
result_2 = mullers_method(func, -5, -2, -1)

print(result_1)
print(func(result_1))
print(result_2)
print(func(result_2))

```

```{r}

```