---
title: Ted Cruz and Term Limits
author: Freddy Drennan
date: January 4, 2019
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Senator Ted Cruz has proposed an [amendment](https://www.cruz.senate.gov/files/documents/Bills/20190103_termlimits.pdf) to the Constitution which would limit the term for a House Representative to 3 terms (6 years) and a senator to 2 terms (12 years). The amendment states that this would go into effect after 7 years of it's acceptance. 

I was curious what this might look like if it was enacted today. The first step is to find a website with the data, scrape it, and then clean it up. We'll use `rvest` to scrape the data. 


```{r}
library(rvest)
library(lubridate)
library(tidyverse)
library(stringr)
```

We're going to scrape from [this](https://ballotpedia.org/List_of_current_members_of_the_U.S._Congress) site. 

```{r}
url <- "https://ballotpedia.org/List_of_current_members_of_the_U.S._Congress"
site <- read_html(url) %>% 
  html_table(fill = TRUE)
```

We see that the relevant data is in the first two tables. 
```{r}
senate <- site %>% 
  .[[1]]

house <- site %>% 
  .[[2]]

DT::datatable(senate)
DT::datatable(house)
```

There are a few issues with the data. First, the dates are strings and we need to convert them to an actual date format in R for plotting. I also prefer my column names to be lower case and without spaces. 

The function below creates a binary indicator stating whether the person is Republican or not. I'm only curious about Democrats and Republicans, so I filter out Independents. Finally, I create a column `years_in_office` which is self explanatory.

```{r}
clean_table <- function(table) {
  table %>% 
    mutate(
      year_assumed = case_when(
        str_detect(`Date assumed office`, "\\-") ~ ymd(`Date assumed office`),
        TRUE ~ ymd(paste0(`Date assumed office`, "-01-01")))
    ) %>% 
    transmute(name = `Officeholder name`,
              party = `Party affiliation`,
              year_assumed,
              republican = if_else(party == "Republican Party", 1, 0),
              years_in_office = year(Sys.time()) - year(year_assumed)) %>% 
    arrange(year_assumed)
}
```

We clean the data and the add a key stating which branch of government the data belongs to.
```{r}
house_clean = clean_table(house) %>% mutate(key = "house")
senate_clean = clean_table(senate) %>% mutate(key = "senate")
```

Next, we select Democrats and Republicans and create a column called `too_long` which indicates wether the individual in question would technically be out of office at this point if the legislation was enforced today.

```{r}
washington = bind_rows(house_clean, senate_clean) %>% 
  filter(party %in% c("Democratic Party", "Republican Party")) %>% 
  mutate(too_long = case_when(
    key == "senate" ~ 12 < years_in_office,
    key == "house"  ~ 6 < years_in_office
  ))

DT::datatable(washington)
```

Next we can plot year year the person came into office and wether he or she is a Republican. In looking at the plot we see generally that Democrats tend to have more members what have been around for a while in the House and vice-verse in the Senate. 

```{r}
washington %>% 
  ggplot() +
  aes(x = year_assumed, y = republican, colour = party) +
  geom_point() +
  facet_grid(key ~.) +
  theme(axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  geom_jitter(width = .1, height = .05)
```

Now let's look at the numbers. If we were to enact the legislation today, how many members from each party will need to go? Overall, it's about even between parties.
```{r}
washington %>% 
  filter(too_long) %>% 
  group_by(party) %>% 
  summarise(n = n())
```

What about by party, by branch? A few more Democrats in the House vs. a few more in the Senate. Overall, I don't get the impression that there is a huge political motivation behind the legislation.
```{r}
washington %>% 
  filter(too_long) %>% 
  group_by(party, key) %>% 
  summarise(n = n())
```